---
title: "Spatial data analyis of Parana State's average rainfall"
author: "Suhas Shastry"
date: "March 10, 2019"
output: pdf_document
---

```{r setup, include=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,fig.pos='h')
library(geoR)
library(gstat)
library(sf)
library(raster)
library(ggplot2)
data(parana)
library(spatstat)
library(raster)
library(gridExtra)
```

\subsection*{Introduction}
Average rainfall over different years for the period May-June (dry-season) was collected at 143 recording stations located at different places in Parana State, Brasil. Below is the graphical analysis of the rainfall data. From data vs X coord scatter plot, we can infer that as we move to the eastern parts of state, average rainfall decreases. From Y coord vs data scatter plot, we can infer that as we move to the northern parts of state, average rainfall decreases. Rainfall is heavier in south-western parts compared to north-eastern parts.

```{r fig.height=3.5,include=FALSE,warning=FALSE}
model <- lm(data~east*north,parana)
s <- summary(model)
df <- data.frame(x = parana$coords[,1],y = parana$coords[,2],data = model$residuals)
```

```{r echo=FALSE,fig.height=3,fig.width=10}
df1 <- parana
df1$data <- model$residuals
#par(mar = c(1,1,1,1) + 0.1)
p1 <- ggplot()+geom_point(mapping=aes(x=parana$coords[,1],y=parana$coords[,2]),size = parana$data/100)+
  geom_polygon(mapping=aes(x=parana$borders[,1],y=parana$borders[,2]),color='black',fill=NA)+theme_bw()+xlab("East")+ylab("North")
p2 <- ggplot()+geom_point(mapping=aes(x=parana$coords[,1],y=parana$data))+theme_bw()+xlab("East")+ylab("Rainfall")
p3 <- ggplot()+geom_point(mapping=aes(x=parana$data,y=parana$coords[,2]))+theme_bw()+xlab("Rainfall")+ylab("North")
grid.arrange(p1,p2,p3, ncol=3)
```

\subsection*{Regression Analysis}
A linear regression model was fit with rainfall as response and distance along east direction, distance along north direction and their interaction as explanatory variables.
$$rainfall = \beta_0 + \beta_1 x_{E}+\beta_2x_N +\beta_3(x_E*x_N) +\epsilon $$
Below are the co-efficients and the p values indicating their significance.

```{r echo=FALSE}
knitr::kable(s$coefficients)
```

As inferred from the plot, east and north co-efficients are significant and negative. Surprisingly interaction effect is positive yet negligibly small.

\subsection*{Spatial Analysis}
As location paramaters are significant predictors of rainfall, exponential covaraince function of the form given below was fit to the spatial data.
$$C(\textbf{x},\textbf{x}+h) = \tau^2 + \sigma^2e^{-\lambda h}  $$

```{r include=FALSE,warning=FALSE}
h <- seq(0,300,10)
variog_data <- variog(as.geodata(df), uvec = h)
v_fit = variofit(variog_data, weights = "npairs",
         cov.model= "exponential",ini.cov.pars = c(1,1))

ml <- likfit(as.geodata(df),
fix.kappa=T, kappa=.5, lik.method="REML",
fix.nugget=FALSE, cov.model="exponential", nugget=.15, 
ini.cov.pars=c(.05,1))
coefs <- data.frame(Coefficients = c('tau-sq','sigma-sq','lambda'),
                    Estimates = round(c(391.56,793.17,1/199.359),2))
```

```{r echo=FALSE}
knitr::kable(coefs)
```

The estimators for the covariance function are calculated using MLE which are given in the above table. Below is the plot of exponential variogram model fitted on the spatial variance.

```{r echo=FALSE,fig.height=2.5,fig.width=5,fig.align='center',fig.pos='h'}
par(mar = c(2,2,1,1) + 0.1)
plot(variog_data,main="Exponential variogram model")
lines.variomodel(cov.model = "exponential", cov.pars = ml$cov.pars, nugget 
= ml$nugget, max.dist = max(h), kappa=ml$kappa)
```

```{r include=FALSE,warning=FALSE}
k.c <- krige.control(obj.model = ml)
o.c <- output.control(simulations.predictive = 1000,n.predictive = 10000,
                      quantile = c(0.1,0.25,0.5,0.75,0.9),threshold = 350)
parana$grid <- pred_grid(parana$borders, by =15)
kc <- krige.conv(parana,loc=parana$grid,krige = k.c,output = o.c)
```

This exponential covaraince function was used for kriging across the state. Predicted values of average rainfall along the state is plotted below.

```{r echo=FALSE,fig.height=3,fig.width=5,fig.align='center'}
par(mar = c(2,2,1,1) + 0.1)
image(kc,main="Kriging of rainfall across the state",col=terrain.colors(20),
      x.leg=c(500,750),y.leg=c(0,50))
```

\subsection*{Conclusion}
Dataset has 143 entries of average rainfall across the state. Initally a linear regression model helped us to infer that location parameters are significant predictors of rainfall. This analysis was extended by employing spatial statistics. Exponential variogram was a good fit to explain the spatial covariance. Finally kriging was applied to predict the rainfall across the state.